---
jupyter:
  title: "Resys_Project"
  kernelspec:
    display_name: R
    language: R
    name: ir
  language_info:
    codemirror_mode: r
    file_extension: .r
    mimetype: text/x-r-source
    name: R
    pygments_lexer: r
    version: 3.6.1
  nbrmd_format_version: '1.0'
  output: html_Resys_Project_notebook
---

Hi

# Introduction
The project starts off simply with the installation of the required packages.
First, we make sure BiocManager is installed, which is a necessity for pretty much every subsequent bioinfo package we will be using next.
Then, we install RTN (for inferring regulatory network), snow (for parallel processing in RTN), and finally RedeR for visualization.

For preprocessing, we install Biobase and GEOquery for base functions in Bioconductor and accessing GEO data respectively.

```{r echo=TRUE, eval=FALSE}
# Installation of required packages, only run if not installed yet

# RTN for inferring the regulatory network based on the downloaded gene expression data
# snow for parallel processing in RTN
# RedeR for better visualization of inferred network

if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

# For main code
BiocManager::install("RTN")
BiocManager::install("snow")
BiocManager::install("RedeR")

# For preprocessing
BiocManager::install("Biobase")
BiocManager::install("GEOquery")

```

Activate libraries (and open documentation for RTN and RedeR)

```{r echo=TRUE, eval=FALSE}
# Load required libraries and open vignettes
# (A vignette is a long-form guide to your package.)
library(igraph)
library(RTN)
library(snow)
library(RedeR)
browseVignettes('RTN')
browseVignettes('RedeR')
# Note: no vignettes for snow, use snow.pdf found at https://cran.r-project.org/web/packages/snow/index.html 

# For preprocessing
library(Biobase)
library(GEOquery)

```

# Preprocessing, as provided in the Slack channel
We get the necessary files remotely and put them inside data objects in our environment.
We do some stuff on the data(?) and throw into gset to run exp and log functions (?)

```{r echo=TRUE, eval=FALSE}
# load series and platform data from GEO

Sys.setenv("VROOM_CONNECTION_SIZE" = 131072 * 2) # to prevent size connection error on getGEO
readr::local_edition(1) # to prevent row.names error on getGEO

gset <- getGEO("GSE21257", GSEMatrix =TRUE, AnnotGPL=FALSE) # get GEO data from notebook
if (length(gset) > 1) idx <- grep("GPL10295", attr(gset, "names")) else idx <- 1 # 
gset <- gset[[idx]]

# log2 transform
ex <- exprs(gset)
exprs(gset) <- log2(ex)
```

# Get the gene names:
(Download full table here https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GPL10295)
Or simply present in the MINiML files as "GPL10295-tbl-1.txt", which we need to add the header to from this link: https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?view=data&acc=GPL10295&id=10175&db=GeoDb_blob44

```{r echo=TRUE, eval=FALSE}

annotation <- read.csv(file = 'GPL10295-tbl-1.txt', skip = 7, sep = '\t') 
gexpIDs <- annotation[,c('ID', 'Illumina_Gene', 'Symbol')]

```


# Start RTN:
Note: the below cell won't run yet, since we don't have dt4rtn (and not sure what it is yet?)

```{r echo=TRUE, eval=FALSE}
data(dt4rtn)

tfs <- dt4rtn$tfs
names(tfs) -> bkp
tfs <- annotation$ID[match(names(tfs), annotation$Symbol)]
tfs <- as.character(tfs)
names(tfs) <- bkp

tfs <- tfs[-which(is.na(tfs))]

rtni <- tni.constructor(expData=ex, regulatoryElements=tfs, rowAnnotation=gexpIDs, cvfilter=T)

```



